Reproducible development environments with rix

### 2024-02-09 paper.qmd
Update paper.qmd

> The idea is to always deploy component closures: if we deploy a
> component, then we must also deploy its dependencies, their dependencies, and so on. That is, we must
> always deploy a set of components that is closed under
> the ‘‘depends on’’ relation. Since closures are selfcontained, they are the units of complete software
> deployment. After all, if a set of components is not
> closed, it is not safe to deploy, since using them might
> cause other components to be referenced that are missing on the target system.

@@

But why is installing software with [Nix]{.proglang} reproducible? Because the
whole set of [Nix]{.pkg} expressions is hosted on Github, it is possible to
use a specific commit of `nixpkgs` to ensure reproducibility of our project.
Using a specific commit of `nixpkgs` is called *pinning a revision (or commit)*.
Pinning a revision ensures that every package that Nix installs will always be
at exactly the same versions, regardless of when in the future the packages get
built. This is because the [Nix]{.pkgs} expressions that get downloaded when
pinning are the expressions as they were at that commit of `nixpkgs`. This is
not the only reason [Nix]{.pkg} is very useful for reproducibility, but the
[Nix]{.pkg} package manager itself was designed with reproducibility in
mind. [Nix]{.pkg} is functional package manager, meaning that it uses
concepts from functional programming to build software: no-side effects or
global variables are allowed, and the output is always the same if given the
same inputs, regardless of where or when a package gets built. [Nix]{.pkg} 
also allows for multiple versions (or variants in the [Nix]{.pkg} jargon)
to be installed at the same time on the same system. Each variant of a package
has a unique identifier, so that it is possible to have two versions of [R]{.proglang}
installed and use the correct one where adequate.
We refer to @dolstra2004nix for a more technical discussion on [Nix]{.pkg}.

>>>>

> The idea is to always deploy component closures: if we deploy a
> component, then we must also deploy its dependencies, their dependencies, and so on. That is, we must
> always deploy a set of components that is closed under
> the ''depends on'' relation. Since closures are selfcontained, they are the units of complete software
> deployment. After all, if a set of components is not
> closed, it is not safe to deploy, since using them might
> cause other components to be referenced that are missing on the target system.

@@

But why is installing software with [Nix]{.proglang} reproducible? Because the
whole set of [Nix]{.pkg} expressions is hosted on Github, it is possible to
use a specific commit of `nixpkgs` to ensure reproducibility of our project.
Using a specific commit of `nixpkgs` is called *pinning a revision (or commit)*.
Pinning a revision ensures that every package that Nix installs will always be
at exactly the same versions, regardless of when in the future the packages get
built. This is because the [Nix]{.pkgs} expressions that get downloaded when
pinning are the expressions as they were at that commit of `nixpkgs`. This is
not the only reason [Nix]{.pkg} is very useful for reproducibility, but the
[Nix]{.pkg} package manager itself was designed with reproducibility in
mind. [Nix]{.pkg} is functional package manager, meaning that it uses
concepts from functional programming to build software: no-side effects or
global variables are allowed, and the output is always the same if given the
same inputs, regardless of where or when a package gets built (an important
side-note is in order here: while the functional aspect of [Nix]{.pkg} is quite
useful to ensure reproducibility, it can lead to situations where writing the
required [Nix]{.proglang} expression to build it a can be much more difficult than usual: this is especially true for
packages that must download assets off the internet during installation, such as 
several BioConductor packages like [musData]{.pkg} for example. But this is only
a concern for the packagers, not users of the package that simply wish to install
and use the package). [Nix]{.pkg} also allows for multiple versions (or variants in the [Nix]{.pkg} jargon)
to be installed at the same time on the same system. Each variant of a package
has a unique identifier, so that it is possible to have two versions of [R]{.proglang}
installed and use the correct one where adequate.
We refer to @dolstra2004nix for a more technical discussion on [Nix]{.pkg}.
